### 2023-11-27 main.tex
giuygiuy

 %Università degli Studi di Modena e Reggio Emilia
(a) University of Parma, \\
Department of ....,\\
Via ...., .... Parma, Italy\\

@@

(b) University of Modena and Reggio Emilia, \\
Department of Physics, Informatics and Mathematics,\\
Via G. Campi 213/b, 41125 Modena, Italy\\ 
\par\endgroup
\vspace{2em}

 Version: \today
\end{center}
%

@@

\begin{abstract}
\noindent
...
\end{abstract}

@@

%The case of $\twoj=1$ distinguishes itself from other values of $\twoj$ as in this case the model can be mapped to an integrable higher rank $gl(N)$ Heisenberg XXX spin chain with open (non-diagonal) boundaries. Furthermore the matrix product ansatz \cite{derrida1993exact}  is applicable, see \cite{Crampe:2014aoa,vanicat2017exact}.

@@

Consider the Lie algebra $gl(N)$ with generators denoted by $\EE_{ab}$ with $a,b\in \{1,\ldots,N\}$ and commutation relations
\begin{equation}\label{eq:comgl}
\left[\EE_{ab},\EE_{cd}\right]=\EE_{ad}\delta_{cb}-\EE_{cb}\delta_{ad}\qquad \forall a,b\in \{1,\ldots,N\}\,.
\end{equation}
The finite-dimensional representations are labelled by partitions $\lambda=(\lambda_1,\lambda_2,\ldots,\lambda_N)$ of $\nu$ with  $\lambda_i\geq \lambda_{i+1}$,  $\lambda_i\in \mathbb{N}$ and $\sum_{i=1}^N \lambda_i = \nu\in\mathbb{N}$.  
We are interested in the {\em symmetric} finite-dimensional representations with 
\begin{equation}\label{eq:dynkin}
    \lambda=(\twoj,0,\ldots,0) \;.
    %\qquad\text{where}\qquad \twoj\in\mathbb{N}\,.
\end{equation} 
The dimension $M_\twoj$ of this symmetric representations is given by the combination of $N$ objects in $\twoj$ positions with repetition, namely
\begin{equation}
	M_\twoj= \frac{(N+\twoj-1)!}{\twoj  !(N-1)!}\dt
\end{equation} 
The generators of the symmetric representations will be denoted by $E_{ab}$.
A basis of the vector space $\mathbb{C}^{M_\twoj}$ are the column vectors denoted by
\begin{equation}
  |n\rangle=  |n_{1},\ldots,n_{N}\rangle,\quad \text{with}\quad n_{i}\in\mathbb{N}_{0}\quad \text{sucht that}\quad \sum_{i=1}^{N}n_{i}=\nu\,.
\end{equation}
%where $\sum_{a=1}^{N}n_{i}=\nu$. 
The basis vectors satisfy the orthogonality relation
\begin{equation}\label{ortho}
   \langle m|n \rangle =\langle m_{1},\ldots,m_{N}|n_{1},\ldots,n_{N}\rangle=\prod_{a=1}^{N}\delta_{m_{a},n_{a}}\co
\end{equation}
where  $ \langle m_{1},\ldots,m_{N}|$ is the row vectors obtained by transposing $|m\rangle=|m_{1},\ldots,m_{N}\rangle$ and $\delta_{m_{a},n_{a}}$ is the Kronecker delta.

@@

The process with generator \eqref{Generator} can be described in terms of $gl(N)$ Lie algebra generators. The state space \eqref{stateSpace} is given by the $|V|$-fold tensor product of the vector space with basis elements $|n^x\rangle$ at a given site. Namely,

@@

We can write the edge Hamiltonian \eqref{edgeHamiltonian} as a function of the coproduct of the quadratic Casimir of $gl(N)$

@@

that acts diagonally as $C|\bm{n}\rangle=\twoj(\twoj+N)|\bm{n}\rangle$ and belogs to the center of $gl(N)$ (i.e. it commutes with all the algebra elements).  

@@

\Delta:gl(N)&\to gl(N)\otimes gl(N)\\

@@

We can write this Hamiltonian in function of the coproduct of the second Casimir
\begin{equation}
    C_{2}=\sum_{a,b=1}^{N}E_{ab}E_{ba}
\end{equation}
It reads
\begin{equation}
	\mathcal{H}_{x,y}=\left\{\frac{1}{2}\Delta^{xy}(C_{2})-\twoj(2\twoj+N)\frac{1}{2}\mathbbm{1}^{x}\otimes\mathbbm{1}^{y}\right\}
\end{equation}
Here we introduced the standard coproduct
\begin{equation}
\begin{split}
\Delta:gl(N)&\to gl(N)\otimes gl(N)\\
E_{ab}&\to E_{ab}\otimes \mathbbm{1}+\mathbbm{1}\otimes E_{ab}
\end{split}
\end{equation} acting on sites $x,y$ and used that $C_{2}|n\rangle=\twoj(\twoj+N)|n\rangle$. 
\end{comment}

@@

We will denote the element of the representation of the $gl(N)$  Lie algebra with $\nu=1$ by lowercase letters, i.e. $(e_{ab})_{a,b\in\{1,\ldots,N\}}$, to highlight that we are in the first fundamental representation of the algebra.

@@

We first look for the similarity between $\mathcal{H}_{x,y}$ and its transposed. Using \eqref{transpositionPropertyR} we obtain  
\begin{equation}\label{transpositionPropertyH}
    \mathcal{H}_{x,y}^{T}=\left(R_{x}R_{y}\right)\mathcal{H}_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the similarity transformation between $\mathcal{H}_{x,y}$ and its transposed. \\
We now look for a symmetry of $\mathcal{H}_{x,y}$, i.e. a matrix $S_{x,y}$ of the same dimension such that it satisfies 
\begin{equation}
	\mathcal{H}_{x,y}S_{x,y}=S_{x,y}\mathcal{H}_{x,y}
	%\quad \Leftrightarrow \quad [\mathcal{H}_{x,y},S_{x,y}]=0
	\dt
\end{equation}
Using \eqref{hamiltonianCasimir}, we observe that $\mathcal{H}_{x,y}$ is proportional to the coproduct of the second Casimir, up to a diagonal term. Therefore, it is enough to look for a symmetry of $\Delta (C)$. Using the bilinearity of the coproduct operator and the fact that $C$ belongs to the centre of the Lie algebra, it is easy to show that for any matrix $E\in gl(N)$ the following holds:
\begin{equation}
 \text{if } \quad	[C,E]=0\qquad  \text{ then }\qquad \left[\Delta (C),\Delta(E)) \right]=0\dt
\end{equation}
We chose 
\begin{equation}
	E=\sum_{a=1}^{N-1}E_{aN}
\end{equation}
and, for a fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
	S_{x,y}=\exp{(\Delta_{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta_{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. As a consequence, this operator $S_{x,y}$ satisfies
\begin{equation}\label{symmetryH}
	\left[S_{x,y},\mathcal{H}_{x,y}\right]=0
\end{equation}
i.e. it is a symmetry of $\mathcal{H}_{x,y}$. Exploiting these considerations we may write

\begin{comment}
As already pointed out, $\mathcal{H}_{x,y}$ is a linear function of $\Delta(C_{2})$. This implies that symmetries of  $\Delta(C_{2})$ are also symmetries of  $\mathcal{H}_{x,y}$. Moreover, the coproduct is a Lie algebra homomorphism, i.e. $\forall A,B\in gl(N)$
\begin{equation}\label{symmetryCoproduct}
    [A,B]=0\quad \Longrightarrow\quad \left[\Delta(A),\Delta(B)\right]=0
\end{equation}
This tells us that it is enough to find a symmetry of $C_{2}$. This element belongs to the center of $gl(N)$, and therefore it commutes with all the element of this algebra. Consider
\begin{equation}\label{equationE}
    E=\sum_{a=2}^{N}E_{a1}
\end{equation}
To obtain a product structure of the elements of the duality matrix, we introduce
\begin{equation}
    S=\exp{(\Delta(E))}
\end{equation}
Finally, for fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
    S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. This operator $S^{x,y}$ satisfies
\begin{equation}\label{symmetryH}
    \left[S_{x,y},\mathcal{H}_{x,y}\right]=0
\end{equation}
i.e. $S_{x,y}$ is a symmetry of $\mathcal{H}_{x,y}$. \\ Exploiting these considerations we write
\end{comment}
\begin{equation}
    \begin{split}
        \mathcal{H}_{x,y}^{T}D&=(R_{x}R_{y})\mathcal{H}_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{u(y)}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{u(z)}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{u(y)}\right)\mathcal{H}_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{u(z)}\right)
        \\&=
        D\mathcal{H}_{x,y}
    \end{split}
\end{equation}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 %\textbf{Reamrk}: it is important to notice that the elements of the duality matrix \eqref{dualityElements} are well defined only if at each site $x\in V$ and for every species $a\in \{2,\ldots,N\}$ the number of dual particles $r_{x}^{a}$ is lower or equal than the number of original particles $n_{a}^{x}$. This implies that the dynamics of the dual process is simpler, because described by a lower number of particles. By the way, it is possible to perform computations similar to the one made in this proof choosing instead of  \eqref{equationE} {\color{red} $E=\sum_{a=2}^{N}E_{1a}$ or $E=\sum_{a=2}^{N}E_{aa}$.}
 %{\color{blue} That's confusing. What is E?}

% However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of boundary duality \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $u(x)$. % that we connect to every site $x\in V$ of the graph $G$. \\
 Considering $A,B\in gl(N)$, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-B)}A\exp{(B)}=A-\left[B,A\right]+\frac{1}{2!}\left[B,\left[B,A\right]\right]-\frac{1}{3!}\left[B,\left[B,\left[B,A\right]\right]\right]+\ldots
 \end{equation}
In the following we need the application of this with $B=\sum_{a=1}^{N-1}E_{aN}$ defined in \eqref{EquationEx} and with $A$ equal to some of the generator $E_{cd}$ of the Lie algebra $gl(N)$ with $c,d\in\{1,\ldots,N\}$. Therefore we compute 
\begin{align}
	\left[B,A\right]=\sum_{a=1}^{N-1}\left[E_{aN},E_{cd}\right]=\delta_{cN}\sum_{a=1}^{N-1}E_{ad}-(1-\delta_{dN})E_{cN}
\end{align}
and 
\begin{align}
	\left[B\left[B,A\right]\right]=&\left[\sum_{a=1}^{N-1}E_{aN},\delta_{cN}\sum_{b=1}^{N-1}E_{bd}-(1-\delta_{dN})E_{cN}\right]\nonumber\\=&
	\sum_{a=1}^{N-1}\left(\delta_{cN}\sum_{b=1}^{N-1}\left[E_{aN},E_{bd}\right]-(1-\delta_{dN})\left[E_{aN},E_{cN}\right]\right)\nonumber\\=&
	\sum_{a=1}^{N-1}\left(-\delta_{cN}\delta_{ad}\sum_{b=1}^{N-1}E_{bN}-(1-\delta_{dN})\delta_{cN}E_{aN}\right)\nonumber\\=&
	-\left(\delta_{cN}(1-\delta_{dN})+(1-\delta_{dN})\delta_{cN}\right)\sum_{b=1}^{N-1}E_{bN}\nonumber
	\\=&
	-2\delta_{cN}(1-\delta_{dN})\sum_{b=1}^{N-1}E_{bN}\dt
\end{align}
From the third commutator on we always obtain zero, i.e. $[B,[B,[B,A]]]=0$. All in all we have that 
 \begin{align}\label{HT-BA}
 	\exp{\left(-\sum_{a=1}^{N-1}E_{aN}\right)}E_{cd}\exp{\left(\sum_{a=1}^{N-1}E_{aN}\right)}=E_{cd}-\delta_{cN}\sum_{a=1}^{N-1}E_{ad}+(1-\delta_{dN})E_{cN}-\delta_{cN}(1-\delta_{dN})\sum_{b=1}^{N-1}E_{bN}\dt
 \end{align}
\begin{comment}
\begin{enumerate}
    \item for $A=E_{b N}$ with $b\in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT_El1}
        \exp{(-E)}E_{b N}\exp{(E)}=E_{b N}
    \end{equation}
    because $E_{b N}$ commutes with $E$
    \item for $A=E_{bb}$ with $b \in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT_Ell}
        \exp{(-E)}E_{b b}\exp{(E)}=E_{b b}+E_{b N}
    \end{equation}
Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
       \left[E, E_{bb}\right]=\sum_{a=1}^{N-1}\left(E_{ab}\delta_{b N}-E_{b N}\delta_{ab}\right)=-E_{bN}
    \end{equation}
    Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT_Ell}.
    \item for $A=E_{Nb}$ with $b\in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT-E1l}
        \exp{(-E)}E_{N a}\exp{(E)}=E_{Na}+E_{NN}-\sum_{c=1}^{N-1}\left(E_{cN}+E_{ca}\right)
    \end{equation}
  %\com{Why indices have names $b, k, j, a, b$? Fix one set for the generators}

@@

			\item Is it possible to extend the duality result to the boundary-driven case in order to have an absorbing dual process, using the same $gl(N)$ Lie-algebraic structure?

@@

we only prove duality for the bulk generator, since the boundary generators are the same of the multi-species stirring process. In the spirit of Section \ref{sectionDuality}, it is convenient to write the Hamiltonian of the process, that is defined through the basis elements \eqref{actionE} of the $gl(N)$ Lie algebra as 

@@

 The basis element $ (E_{ab})_{a,b\in\{1,\ldots,N\}}$ of $gl(N)$ can also be thought as proper compositions of the rising, and lowering operators of th Heisenberg Lie algebra
  \begin{equation}\label{creationOperators}
	\begin{cases}
	\oad_{a}|n_{1},\ldots,n_{a},\ldots,n_{N}\rangle = |n_{1},\ldots,n_{a}+1,\ldots,n_{N}\rangle\\
	\oa_{a}|n_{1},\ldots,n_{a},\ldots,n_{N}\rangle =n_{a}|n_{1},\ldots,n_{a}-1,\ldots,n_{N}\rangle
	\end{cases}
\end{equation}
with commutaion relations
\begin{equation}
	[\oad_{a},\oa_{b}]=-\mathbbm{1}\,\delta_{ab}
\end{equation}
Therefore, we have that 
\begin{equation}
		E_{ab}=\oad_{a}\oa_{b}\qquad \forall a,b=1,\ldots N
\end{equation} 
fulfil the action on the basis vector \eqref{actionE} and the commutation relation \eqref{eq:comgl} }
\end{comment}

@@

We chose the $gl(N)$ Lie algebra. This is compact. The abstract basis elements are denoted by $E_{a,b}$ with $a,b\in \{1,\ldots,N\}$ that obeys to the Lie bracket relations given by

>>>>

 %Università degli Studi di Modena e Reggio Emilia
(a) University of Parma, \\
Department of Mathematica, Physics, Computer Science,\\
Parco Area delle Scienze, 7/A, 43124 Parma, Italy\\

@@

(b) University of Modena and Reggio Emilia, \\
Department of Physics, Computer Science and Mathematics,\\
Via G. Campi 213/b, 41125 Modena, Italy\\ 
\par\endgroup
\vspace{2em}

% Version: \today
\end{center}
%

@@

\begin{abstract}
\noindent
We study the boundary-driven version of the stirring process with $N-1$ species
on a generic graph $G=(V,\mathcal{E})$ with reservoirs. 
The multi-species stirring process generalizes the symmetric exclusion process, 
which is recovered in the case $N=2$.
We prove the existence of a dual process defined on
an extended graph $\widetilde{G}=(\widetilde{V},\widetilde{\mathcal{E})}$ which 
includes additional extra-sites $\widetilde{V}\setminus V$ where
dual particles gets absorbed in the long-time limit.  
We thus obtain a characterization of the non-equilibrium
steady state of the bounbdary-driven system in terms of the absorption
probabilities of dual particles.
For the case of one-dimensional chains with two reservoirs
at the boundaries and with maximum one particle per site,
we compute these absorption probabilities by relying
on the underlying $\mathfrak{gl}(N)$ symmetry and the
matrix product ansatz. Thus one gets a closed-formula
for (long-ranged) correlations and for the 
non-equilibrium stationary measure.
Extensions beyond this integrable set-up
are also discussed.
\end{abstract}

@@

%The case of $\twoj=1$ distinguishes itself from other values of $\twoj$ as in this case the model can be mapped to an integrable higher rank $\mathfrak{gl}(N)$ Heisenberg XXX spin chain with open (non-diagonal) boundaries. Furthermore the matrix product ansatz \cite{derrida1993exact}  is applicable, see \cite{Crampe:2014aoa,vanicat2017exact}.

@@

Consider the Lie algebra $\mathfrak{gl}(N)$ with generators denoted by $\EE_{ab}$ with $a,b\in \{1,\ldots,N\}$ and commutation relations
\begin{equation}\label{eq:comgl}
\left[\EE_{ab},\EE_{cd}\right]=\EE_{ad}\delta_{cb}-\EE_{cb}\delta_{ad}\qquad \forall a,b\in \{1,\ldots,N\}\,.
\end{equation}
The finite-dimensional representations are labelled by partitions $\lambda=(\lambda_1,\lambda_2,\ldots,\lambda_N)$ of $\nu$ with  $\lambda_i\geq \lambda_{i+1}$,  $\lambda_i\in \mathbb{N}$ and $\sum_{i=1}^N \lambda_i = \nu\in\mathbb{N}$.  
We are interested in the {\em symmetric} finite-dimensional representations with 
\begin{equation}\label{eq:dynkin}
    \lambda=(\twoj,0,\ldots,0) \;.
    %\qquad\text{where}\qquad \twoj\in\mathbb{N}\,.
\end{equation} 
The dimension $M_\twoj$ of this symmetric representations is given by the combination of $N$ objects in $\twoj$ positions with repetition, namely
\begin{equation}
	M_\twoj= \frac{(N+\twoj-1)!}{\twoj  !(N-1)!}\dt
\end{equation} 
The generators of the symmetric representations will be denoted by $E_{ab}$.
A basis of the vector space $\mathbb{C}^{M_\twoj}$ are the column vectors denoted by
\begin{equation}
  |n\rangle=  |n_{1},\ldots,n_{N}\rangle,\quad \text{with}\quad n_{i}\in\mathbb{N}_{0}\quad \text{sucht that}\quad \sum_{i=1}^{N}n_{i}=\nu\,.
\end{equation}
%where $\sum_{a=1}^{N}n_{i}=\nu$. 
The basis vectors satisfy the orthogonality relation
\begin{equation}\label{ortho}
   \langle m|n \rangle =\langle m_{1},\ldots,m_{N}|n_{1},\ldots,n_{N}\rangle=\prod_{a=1}^{N}\delta_{m_{a},n_{a}}\co
\end{equation}
where  $ \langle m_{1},\ldots,m_{N}|$ is the row vectors obtained by transposing $|m\rangle=|m_{1},\ldots,m_{N}\rangle$ and $\delta_{m_{a},n_{a}}$ is the Kronecker delta.

@@

The process with generator \eqref{Generator} can be described in terms of $\mathfrak{gl}(N)$ Lie algebra generators. The state space \eqref{stateSpace} is given by the $|V|$-fold tensor product of the vector space with basis elements $|n^x\rangle$ at a given site. Namely,

@@

We can write the edge Hamiltonian \eqref{edgeHamiltonian} as a function of the coproduct of the quadratic Casimir of $\mathfrak{gl}(N)$

@@

that acts diagonally as $C|\bm{n}\rangle=\twoj(\twoj+N)|\bm{n}\rangle$ and belogs to the center of $\mathfrak{gl}(N)$ (i.e. it commutes with all the algebra elements).  

@@

\Delta:\mathfrak{gl}(N)&\to \mathfrak{gl}(N)\otimes \mathfrak{gl}(N)\\

@@

We can write this Hamiltonian in function of the coproduct of the second Casimir
\begin{equation}
    C_{2}=\sum_{a,b=1}^{N}E_{ab}E_{ba}
\end{equation}
It reads
\begin{equation}
	\mathcal{H}_{x,y}=\left\{\frac{1}{2}\Delta^{xy}(C_{2})-\twoj(2\twoj+N)\frac{1}{2}\mathbbm{1}^{x}\otimes\mathbbm{1}^{y}\right\}
\end{equation}
Here we introduced the standard coproduct
\begin{equation}
\begin{split}
\Delta:\mathfrak{gl}(N)&\to \mathfrak{gl}(N)\otimes \mathfrak{gl}(N)\\
E_{ab}&\to E_{ab}\otimes \mathbbm{1}+\mathbbm{1}\otimes E_{ab}
\end{split}
\end{equation} acting on sites $x,y$ and used that $C_{2}|n\rangle=\twoj(\twoj+N)|n\rangle$. 
\end{comment}

@@

We will denote the element of the representation of the $\mathfrak{gl}(N)$  Lie algebra with $\nu=1$ by lowercase letters, i.e. $(e_{ab})_{a,b\in\{1,\ldots,N\}}$, to highlight that we are in the first fundamental representation of the algebra.

@@

We first look for the similarity between $\mathcal{H}_{x,y}$ and its transposed. Using \eqref{transpositionPropertyR} we obtain  
\begin{equation}\label{transpositionPropertyH}
    \mathcal{H}_{x,y}^{T}=\left(R_{x}R_{y}\right)\mathcal{H}_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the similarity transformation between $\mathcal{H}_{x,y}$ and its transposed. \\
We now look for a symmetry of $\mathcal{H}_{x,y}$, i.e. a matrix $S_{x,y}$ of the same dimension such that it satisfies 
\begin{equation}
	\mathcal{H}_{x,y}S_{x,y}=S_{x,y}\mathcal{H}_{x,y}
	%\quad \Leftrightarrow \quad [\mathcal{H}_{x,y},S_{x,y}]=0
	\dt
\end{equation}
Using \eqref{hamiltonianCasimir}, we observe that $\mathcal{H}_{x,y}$ is proportional to the coproduct of the second Casimir, up to a diagonal term. Therefore, it is enough to look for a symmetry of $\Delta (C)$. Using the bilinearity of the coproduct operator and the fact that $C$ belongs to the centre of the Lie algebra, it is easy to show that for any matrix $E\in \mathfrak{gl}(N)$ the following holds:
\begin{equation}
 \text{if } \quad	[C,E]=0\qquad  \text{ then }\qquad \left[\Delta (C),\Delta(E)) \right]=0\dt
\end{equation}
We chose 
\begin{equation}
	E=\sum_{a=1}^{N-1}E_{aN}
\end{equation}
and, for a fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
	S_{x,y}=\exp{(\Delta_{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta_{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. As a consequence, this operator $S_{x,y}$ satisfies
\begin{equation}\label{symmetryH}
	\left[S_{x,y},\mathcal{H}_{x,y}\right]=0
\end{equation}
i.e. it is a symmetry of $\mathcal{H}_{x,y}$. Exploiting these considerations we may write

\begin{comment}
As already pointed out, $\mathcal{H}_{x,y}$ is a linear function of $\Delta(C_{2})$. This implies that symmetries of  $\Delta(C_{2})$ are also symmetries of  $\mathcal{H}_{x,y}$. Moreover, the coproduct is a Lie algebra homomorphism, i.e. $\forall A,B\in \mathfrak{gl}(N)$
\begin{equation}\label{symmetryCoproduct}
    [A,B]=0\quad \Longrightarrow\quad \left[\Delta(A),\Delta(B)\right]=0
\end{equation}
This tells us that it is enough to find a symmetry of $C_{2}$. This element belongs to the center of $\mathfrak{gl}(N)$, and therefore it commutes with all the element of this algebra. Consider
\begin{equation}\label{equationE}
    E=\sum_{a=2}^{N}E_{a1}
\end{equation}
To obtain a product structure of the elements of the duality matrix, we introduce
\begin{equation}
    S=\exp{(\Delta(E))}
\end{equation}
Finally, for fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
    S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. This operator $S^{x,y}$ satisfies
\begin{equation}\label{symmetryH}
    \left[S_{x,y},\mathcal{H}_{x,y}\right]=0
\end{equation}
i.e. $S_{x,y}$ is a symmetry of $\mathcal{H}_{x,y}$. \\ Exploiting these considerations we write
\end{comment}
\begin{equation}
    \begin{split}
        \mathcal{H}_{x,y}^{T}D&=(R_{x}R_{y})\mathcal{H}_{x,y}(R_{x}R_{y})^{-1}\left(d_{x}\otimes\dd\right)\left(d_{y}\otimes\mathcal{D}_{u(y)}\right)\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{u(z)}\right)
        \\&=\left(R_{x}\exp{(E^{x})}\otimes \dd\right)\left(R_{y}\exp{(E^{y})}\otimes \mathcal{D}_{u(y)}\right)\mathcal{H}_{x,y}\prod_{z\in V\,:\, z\neq x,y}\left(d_{z}\otimes \mathcal{D}_{u(z)}\right)
        \\&=
        D\mathcal{H}_{x,y}
    \end{split}
\end{equation}
 where we used \eqref{transpositionPropertyH} and \eqref{symmetryH} in the second equality. Thus, \eqref{edgeDualRealtion} is proved. 
 %\textbf{Reamrk}: it is important to notice that the elements of the duality matrix \eqref{dualityElements} are well defined only if at each site $x\in V$ and for every species $a\in \{2,\ldots,N\}$ the number of dual particles $r_{x}^{a}$ is lower or equal than the number of original particles $n_{a}^{x}$. This implies that the dynamics of the dual process is simpler, because described by a lower number of particles. By the way, it is possible to perform computations similar to the one made in this proof choosing instead of  \eqref{equationE} {\color{red} $E=\sum_{a=2}^{N}E_{1a}$ or $E=\sum_{a=2}^{N}E_{aa}$.}
 %{\color{blue} That's confusing. What is E?}

% However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of boundary duality \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $u(x)$. % that we connect to every site $x\in V$ of the graph $G$. \\
 Considering $A,B\in \mathfrak{gl}(N)$, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-B)}A\exp{(B)}=A-\left[B,A\right]+\frac{1}{2!}\left[B,\left[B,A\right]\right]-\frac{1}{3!}\left[B,\left[B,\left[B,A\right]\right]\right]+\ldots
 \end{equation}
In the following we need the application of this with $B=\sum_{a=1}^{N-1}E_{aN}$ defined in \eqref{EquationEx} and with $A$ equal to some of the generator $E_{cd}$ of the Lie algebra $\mathfrak{gl}(N)$ with $c,d\in\{1,\ldots,N\}$. Therefore we compute 
\begin{align}
	\left[B,A\right]=\sum_{a=1}^{N-1}\left[E_{aN},E_{cd}\right]=\delta_{cN}\sum_{a=1}^{N-1}E_{ad}-(1-\delta_{dN})E_{cN}
\end{align}
and 
\begin{align}
	\left[B\left[B,A\right]\right]=&\left[\sum_{a=1}^{N-1}E_{aN},\delta_{cN}\sum_{b=1}^{N-1}E_{bd}-(1-\delta_{dN})E_{cN}\right]\nonumber\\=&
	\sum_{a=1}^{N-1}\left(\delta_{cN}\sum_{b=1}^{N-1}\left[E_{aN},E_{bd}\right]-(1-\delta_{dN})\left[E_{aN},E_{cN}\right]\right)\nonumber\\=&
	\sum_{a=1}^{N-1}\left(-\delta_{cN}\delta_{ad}\sum_{b=1}^{N-1}E_{bN}-(1-\delta_{dN})\delta_{cN}E_{aN}\right)\nonumber\\=&
	-\left(\delta_{cN}(1-\delta_{dN})+(1-\delta_{dN})\delta_{cN}\right)\sum_{b=1}^{N-1}E_{bN}\nonumber
	\\=&
	-2\delta_{cN}(1-\delta_{dN})\sum_{b=1}^{N-1}E_{bN}\dt
\end{align}
From the third commutator on we always obtain zero, i.e. $[B,[B,[B,A]]]=0$. All in all we have that 
 \begin{align}\label{HT-BA}
 	\exp{\left(-\sum_{a=1}^{N-1}E_{aN}\right)}E_{cd}\exp{\left(\sum_{a=1}^{N-1}E_{aN}\right)}=E_{cd}-\delta_{cN}\sum_{a=1}^{N-1}E_{ad}+(1-\delta_{dN})E_{cN}-\delta_{cN}(1-\delta_{dN})\sum_{b=1}^{N-1}E_{bN}\dt
 \end{align}
\begin{comment}
\begin{enumerate}
    \item for $A=E_{b N}$ with $b\in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT_El1}
        \exp{(-E)}E_{b N}\exp{(E)}=E_{b N}
    \end{equation}
    because $E_{b N}$ commutes with $E$
    \item for $A=E_{bb}$ with $b \in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT_Ell}
        \exp{(-E)}E_{b b}\exp{(E)}=E_{b b}+E_{b N}
    \end{equation}
Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
       \left[E, E_{bb}\right]=\sum_{a=1}^{N-1}\left(E_{ab}\delta_{b N}-E_{b N}\delta_{ab}\right)=-E_{bN}
    \end{equation}
    Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT_Ell}.
    \item for $A=E_{Nb}$ with $b\in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT-E1l}
        \exp{(-E)}E_{N a}\exp{(E)}=E_{Na}+E_{NN}-\sum_{c=1}^{N-1}\left(E_{cN}+E_{ca}\right)
    \end{equation}
  %\com{Why indices have names $b, k, j, a, b$? Fix one set for the generators}

@@

			\item Is it possible to extend the duality result to the boundary-driven case in order to have an absorbing dual process, using the same $\mathfrak{gl}(N)$ Lie-algebraic structure?

@@

we only prove duality for the bulk generator, since the boundary generators are the same of the multi-species stirring process. In the spirit of Section \ref{sectionDuality}, it is convenient to write the Hamiltonian of the process, that is defined through the basis elements \eqref{actionE} of the $\mathfrak{gl}(N)$ Lie algebra as 

@@

 The basis element $ (E_{ab})_{a,b\in\{1,\ldots,N\}}$ of $\mathfrak{gl}(N)$ can also be thought as proper compositions of the rising, and lowering operators of th Heisenberg Lie algebra
  \begin{equation}\label{creationOperators}
	\begin{cases}
	\oad_{a}|n_{1},\ldots,n_{a},\ldots,n_{N}\rangle = |n_{1},\ldots,n_{a}+1,\ldots,n_{N}\rangle\\
	\oa_{a}|n_{1},\ldots,n_{a},\ldots,n_{N}\rangle =n_{a}|n_{1},\ldots,n_{a}-1,\ldots,n_{N}\rangle
	\end{cases}
\end{equation}
with commutaion relations
\begin{equation}
	[\oad_{a},\oa_{b}]=-\mathbbm{1}\,\delta_{ab}
\end{equation}
Therefore, we have that 
\begin{equation}
		E_{ab}=\oad_{a}\oa_{b}\qquad \forall a,b=1,\ldots N
\end{equation} 
fulfil the action on the basis vector \eqref{actionE} and the commutation relation \eqref{eq:comgl} }
\end{comment}

@@

We chose the $\mathfrak{gl}(N)$ Lie algebra. This is compact. The abstract basis elements are denoted by $E_{a,b}$ with $a,b\in \{1,\ldots,N\}$ that obeys to the Lie bracket relations given by
